---
title: "Data Analysis using R Programming - BUS4028"
author: "Group 08"
date: "2025-03-28"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[C]{\url{https://github.com/musaabshirgar1/Data-Analysis-Using-R/tree/main}}
- \fancyfoot[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this document, we will be working on Data Analysis Assignment for BUS4028. This report performs data analysis using R programming on a climate change dataset. The analysis includes data manipulation, statistical functions, and plotting. 

## Data Loading

### Read CSV file climate_change_indicators.csv
```{r}
df <- read.csv("climate_change_indicators.csv")
```

## Data Exploration

### Print structure of df
```{r}
str(df)
```

### List variables in dataset
```{r}
ls(df)
```

### Print Top 15 Rows of Dataset
```{r}
head(df, 15)
```

## User-Defined Function

### Create a User Defined Function
```{r}
celsius_to_fahreheit <- function(temp) { return ((temp * 9/5) + 32)}
df$F2022_Fahrenheit <- celsius_to_fahreheit(df$F2022)
print(df$F2022_Fahrenheit)
```

## Data Manipulation

### Filtering Rows using subset() where Country = Canada
```{r}
canada_df <- subset(df, ISO3 == "CAN")
head(canada_df)
```

### Identify dependent and Independent Variables and Reshape the data
```{r}
#selected two columns for reshaping i.e. Country and F2022
selected_data <- df[, c("Country", "F2022")]
library(reshape2)
melted_data <- melt(selected_data, id.vars = "Country",
                    variable.name = "Year", value.name = "TemperatureChange")
print(melted_data)
```

### Remove Missing Values from Data Frame
```{r}
df <- na.omit(df)
```

### Identify and Remove Duplicated Data
```{r}
#identify duplicate rows
duplicate_rows <- duplicated(df)
print("Duplicate Rows:")
print(duplicate_rows)

#Remove duplicate rows
unique_df <- unique(df)
print("DataFrame after removing duplicates:")
print(unique_df)
```

### Reorder Multiple Rows in Descending Order
```{r}
ordered_df <- df[order(df$Country, -df$F2022), ]
print("Ordered by Country, then F2022 (descending):")
head(ordered_df)
```

### Renaming Columns
```{r}
colnames(df)[colnames(df) == "F2022"] <- "Temperature_2022"
colnames(df)[colnames(df) == "F2021"] <- "Temperature_2021"
ls(df)
```

### Adding New Variables
```{r}
#adding new column F2022_x_2 which is F2022 multiplied by 2
df$F2022_x_2 <- df$Temperature_2022 * 2
ls(df)
```

### Creating a Training Set
```{r}
#setting seed for reproducibility
set.seed(123)
#creating a vector of random numbers
train_index <- sample(seq_len(nrow(df)), 0.8 * nrow(df))
#create training set
train_set <- df[train_index, ]
#display top 15 rows of train dataset
head(train_set, 15)
```

## Summary Statistics and Statistical Functions

### Summary Statistics
```{r}
print(summary(df))
```

### Statistical Functions
```{r}
#Mean of F2022
mean_F2022 <- mean(df$Temperature_2022) # nolint
print(paste("Mean of Temperature_2022:", mean_F2022))

#Median of F2022
median_F2022 <- median(df$Temperature_2022) # nolint
print(paste("Median of Temperature_2022:", median_F2022))

#Mode of F2022
get_mode <- function(v) {
  unique_val <- unique(v)
  unique_val[which.max(tabulate(match(v, unique_val)))]
}
mode_F2022 <- get_mode(df$Temperature_2022) # nolint
print(paste("Mode of Temperature_2022:", mode_F2022))

#Range of F2022
range_F2022 <- range(df$Temperature_2022) # nolint
print(paste("Range of Temperature_2022:", range_F2022))
```

# Plots and Correlation

### Scatter Plot
```{r}
plot(df$Temperature_2021, df$Temperature_2022,
     main = "Scatter Plot of Temperature_2021 vs Temperature 2022",
     xlab = "Temperature_2021", ylab = "Temperature_2022")
```

### Bar Plot
```{r}
library(ggplot2)
library(dplyr)

# Getting the top 10 countries by Temperature_2022
top_10_countries <- df %>%
  arrange(desc(Temperature_2022)) %>%
  head(10)

# Creating the bar plot with only the top 10
ggplot(top_10_countries, aes(x = factor(Country, levels = Country),
                             y = Temperature_2022)) + 
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Countries by Temperature_2022",
       x = "Country",
       y = "Temperature_2022") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"))
```

### Correlation
```{r}
correlation_F2021_F2022 <- cor(df$Temperature_2021, df$Temperature_2022, use = "complete.obs") # nolint
print(paste("Correlation between Temperature_2021 and Temperature_2022:",
            correlation_F2021_F2022))
```